Constraints on Assembly Bias from Galaxy Clustering

### 2016-06-13 ms.tex
Edited 3.2

We turn now to a discussion of our parameter inference analysis of projected galaxy clustering 
in Decorated HOD models that include a treatment of galaxy assembly bias. In this work, 
we consider only the simplest model of galaxy assembly bias, introducing only two new 
parameters, $A_{\rm cen}$ and $A_{\rm sat}$, that describe the strength of central galaxy 
and satellite galaxy assembly bias respectively. These parameters are limited to values 
of $-1 \le A_{\rm cen,sat} \le 1$, and $A_{\rm cen,sat}=0$ when there is no galaxy assembly bias. 
In this work, we use halo concentration as our secondary halo property, so $A_{\rm cen,sat}=1$ 
($A_{\rm cen,sat}=-1$) means that the mean number of galaxies per halo is maximally 
correlated (anti-correlated) with halo concentration. The model and its implementation 
in {\tt halotools} is discussed further in Section~\ref{subsection:decorated} above and 
in \citet{hearin_etal16}.

Examples of our fits are given in the right-hand panels of Figure~\ref{fig:Mr19samples}, 
Figure~\ref{fig:Mr20samples}, and Figure~\ref{fig:Mr21samples}. The general trend that 
can be gleaned from these figures is that introducing assembly bias improves the ability 
of the predicted two-point functions to match the measured two-point functions across the 
transition form the one-halo (highly nonlinear) to two-halo (nearly linear) regimes near $\rp \sim 2\, h^{-1}{\mathrm{Mpc}}$. 
This is most apparent for the $M_r < -20$ threshold sample shown in Fig.~\ref{fig:Mr20samples}. 
Visually, these differences appear to be small; however, Table~\ref{table:parameters} shows that 
they are statistically important.

The one-dimensional marginalized constraints on all parameters from these analyses 
are given in the lowest row of each luminosity threshold grouping in Table~\ref{table:parameters}. 
In cases where the posterior on a parameter is monotonic within the physical parameter 
range, we quote an upper or lower limit on the parameter. One- and two-dimensional 
visualizations of the posteriors from our analysis are shown in Figure~\ref{fig:Mr19ABtriangle}, 
Figure~\ref{fig:Mr20ABtriangle}, and Figure~\ref{fig:Mr21ABtriangle}.

@@

Table~\ref{table:parameters} and Figures~\ref{fig:Mr19ABtriangle}-\ref{fig:Mr21triangle} 
all make several simple, generic points. Introducing additional parameter 
freedom associated with galaxy assembly bias generally increases the viable parameter 
space, even for the subset of standard HOD parameters. Constraints on the 
standard HOD parameters are generally less restrictive. This is exactly what is 
expected from the introduction of additional parameter freedom.

@@

Focusing attention on the parameters describing galaxy assembly bias, it is 
evident that these parameters are often quite poorly constrained by galaxy 
clustering data. This is important as it implies that galaxy clustering of the precision 
of SDSS DR7 measurements cannot rule out, or strongly restrict galaxy assembly 
bias in many cases. Nonetheless, it is apparent that the presence of assembly bias 
can alter the inferred HOD, or more generally, the inferred relationship between 
galaxies and halo mass. This is most evident for the $M_r < -20$ threshold sample, 
for which there are significant differences in the inferred values of all baseline HOD 
parameters between models with and without galaxy assembly bias. Other threshold 
samples exhibit significant changes particularly for $\alpha$, and to a lesser degree 
for $\log (M_{\rm min})$ and $\sigma_{\log M}$.

Beyond those generic conclusions, a few specific cases are worthy of further examination. 
Consider the $M_r < -20$ sample. The inferred value of $A_{\rm cen} > 0.28$ at 99\% 
confidence. In this case the data strongly prefer $A_{\rm cen} > 0$ and thus strongly prefer 
galaxies to reside in halos of larger concentration at fixed halo mass. This particular threshold 
sample is the most significant outlier in this regard. Nevertheless, there are hints of assembly 
bias in other samples. Satellite galaxies show a marginal preference for occupying halos of 
higher concentration in the $M_r < -19$ threshold sample. The $M_r < -19$ sample exhibits 
weak preference for a positive correlation of galaxy occupation with halo concentration at fixed 
mass for both satellite galaxies and central galaxies. Continuing upward with luminosity, 
the $M_r < -20.5$ sample exhibits a significant preference for central galaxy assembly bias. 
Lastly, there is no preference for either central galaxy or satellite galaxy assembly bias 
for the $M_r < -21$ threshold sample, for which both $A_{\mathrm{cen}}$ and $A_{\mathrm{sat}}$ 
are consistent with zero will within 1$\sigma$. These data suggest that assembly bias 
may be present in the real universe.

>>>>

We turn now to a discussion of our parameter inference analysis of projected galaxy clustering 
in Decorated HOD models that include a treatment of galaxy assembly bias. In this work, 
we consider only the simplest model of galaxy assembly bias, introducing only two new 
parameters, $A_{\rm cen}$ and $A_{\rm sat}$, that describe the strength of central galaxy 
and satellite galaxy assembly bias respectively. These parameters are limited to values 
of $-1 \le A_{\rm cen,sat} \le 1$, and $A_{\rm cen,sat}=0$ when there is no galaxy assembly bias. 
In this work, we use halo concentration as our secondary halo property, so $A_{\rm cen,sat}=1$ 
($A_{\rm cen,sat}=-1$) means that the mean number of galaxies per halo is maximally 
correlated (anti-correlated) with halo concentration. The model and its implementation 
in {\tt halotools} is discussed further in Section~\ref{subsection:decorated} above and 
in more detail in \citet{hearin_etal16}.

Examples of our fits are given in the right-hand panels of Figure~\ref{fig:Mr19samples}, 
Figure~\ref{fig:Mr20samples}, and Figure~\ref{fig:Mr21samples}. The general trend that 
can be gleaned from these figures is that introducing assembly bias improves the ability 
of the predicted two-point functions to match the measured two-point functions across the 
transition from the one-halo (highly nonlinear) to two-halo (nearly linear) regimes near $\rp \sim 2\, h^{-1}{\mathrm{Mpc}}$. 
This is most apparent for the $M_r < -20$ threshold sample shown in Fig.~\ref{fig:Mr20samples}. 
Visually, these differences appear to be small; however, Table~\ref{table:parameters} shows that 
they are statistically important.

The one-dimensional marginalized constraints on all parameters from these analyses 
are given in the lowest row of each luminosity threshold grouping in Table~\ref{table:parameters}. 
In cases where the posterior on a parameter is monotonic within the physical parameter 
range, we quote an upper or lower limit on the parameter. Examples of one- and two-dimensional 
visualizations of the posteriors from our analyses are shown in Figure~\ref{fig:Mr19ABtriangle}, 
Figure~\ref{fig:Mr20ABtriangle}, and Figure~\ref{fig:Mr21ABtriangle} for the 
$M_r<-19$, $M_r<-20$, and $M_r<-21$ samples respectively.

@@

Table~\ref{table:parameters} and Figures~\ref{fig:Mr19ABtriangle}--\ref{fig:Mr21ABtriangle} 
all make several simple, generic points. Introducing additional parameter 
freedom associated with galaxy assembly bias generally increases the viable parameter 
space, even for the subset of standard HOD parameters. Constraints on the 
standard HOD parameters are generally less restrictive. This is exactly what is 
expected from the introduction of additional parameter freedom.

@@

Focusing attention on the parameters describing galaxy assembly bias, it is 
evident that these parameters are often quite poorly constrained by galaxy 
clustering data. This is important as it implies that galaxy clustering of the precision 
of SDSS DR7 measurements cannot rule out, or strongly restrict galaxy assembly 
bias in many cases. This is the case, for example, for the $M_r<-21$ sample with posteriors 
shown in Fig.~\ref{fig:Mr21ABtriangle}. In this case, both $A_{\rm cen}$ and 
$A_{\rm sat}$ have posteriors that peak near zero, but are very broad. 
Nonetheless, it is apparent that the presence of galaxy assembly bias 
can alter the inferred HOD, or more generally, the inferred relationship between 
galaxies and halo mass. This is most evident for the $M_r < -20$ threshold sample, 
for which there are significant differences in the inferred values of all baseline HOD 
parameters between models with and without galaxy assembly bias. Other threshold 
samples exhibit significant differences particularly for $\alpha$, and to a lesser degree 
for $\log (M_{\rm min})$ and $\sigma_{\log M}$.

Beyond those generic conclusions, a few specific cases are worthy of further examination. 
Consider the $M_r < -20$ sample. The inferred value of $A_{\rm cen} > 0.28$ at 99\% 
confidence. In this case the data strongly prefer $A_{\rm cen} > 0$ and thus strongly prefer 
galaxies to reside in halos of larger concentration at fixed halo mass. Figure~\ref{fig:directhodcomparison} 
compares the inferred HODs for the $M_r<-20$ sample in a standard HOD analysis with the 
inferred HOD from our decorated HOD analysis that includes parameters to describe 
assembly bias. This is an explicit example of the degree to which the inferred relationship 
between galaxies and halos can be altered by assembly bias.

Though the evidence for assembly bias is the most significant for the $M_r<-20$ sample, there are hints of assembly 
bias in other samples. Satellite galaxies show a marginal preference for occupying halos of 
higher concentration in the $M_r < -19$ threshold sample. The $M_r < -19$ sample exhibits 
weak preference for a positive correlation of galaxy occupation with halo concentration at fixed 
mass for both satellite galaxies and central galaxies. Continuing upward with luminosity, 
the $M_r < -20.5$ sample exhibits a significant preference for central galaxy assembly bias. 
Lastly, there is no preference for either central galaxy or satellite galaxy assembly bias 
for the $M_r < -21$ threshold sample, for which both $A_{\mathrm{cen}}$ and $A_{\mathrm{sat}}$ 
are consistent with zero will within 1$\sigma$. These data suggest that assembly bias 
may be present in the real universe.
